Time stability in human interaction networks

### 2015-05-27 paper.tex
finished review of article regarding fb tt and pbr networks

\begin{abstract}
	In this study, we demonstrate a remarkably stable activity in networks derived from e-mail lists. The activity along time and topology evolution were investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparying with the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. A clear distinction could be made of networks with few, stable agents, which possess a large number of threads per message, from others with many agents exhibiting intermittent activity and lower number of threads per message. Also important was the observation that the activity along time, in terms of seconds, minutes, hours, days and months, is practically the same for all lists. These results are coherent with literature and were found also in the analysis of other networks from other emails lists and from Facebook and Twitter. Therefore, we infer that the properties observed may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{newmanBook}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{gmane,barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed ~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or functionality, such as accessibility or community detection~\cite{access,newmanModularity}.

In this paper we analyze the evolution of human interaction networks, by considering interaction in email lists as their representative. Using a timeline of activity snapshots with a constant number of contiguous messages in email lists, we found a remarkable stability for several of the network properties. Because this stability was shared by all email lists, is coherent with literature, and has been observed in some networks from Twitter and Facebook, we advocate that some of the conclusions might be valid for more general classes of interaction networks. In particular, this allows us to discuss typologies in the context of such networks, in an attempt to bridge the gap between approaches based solely on data analysis (i.e. from a hard sciences perspective) and those relevant to the social sciences. This is important insofar as typologies are the canon of scientific literature for classification of human agents~\cite{typCanon}.

The paper is organized as follows. Section~\ref{sec:related} describes related work, while details of the data and methods of analysis are given in  Section~\ref{sec:data} and Section~\ref{sec:carac}. Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions.

@@

We analyzed many email lists (and data from Facebook and Twitter) but selected only four in order to make a thorough analysis, from which general properties can be inferred. These lists, selected as representing both a diverse set and ordinary lists, are:
\begin{itemize}
	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries. English is the language used the most. Abbreviated as LAU from now on.
	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants from different countries, and English is the language used the most. A more technical and less active version of LAU. Abbreviated LAD from now on.
	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries. English is the language used the most. Abbreviated as CPP from now on.
	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests. Portuguese is the most used language, although Spanish and English are also incident. Abbreviated MET from now on.
\end{itemize}

@@

A systematic study of the activity of participants belonging to the three distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices, while the network structure was governed by the intermediary vertices. Because these properties were shared by all email lists and were time-independent, are coherent with literature, and was also found in networks from Facebook and Twitter, we may consider the classification into primitive Erd\"os sectors as leading to a typology which bridges exact sciences, with objective procedures for classifying agents, with human sciences, where there is a legacy in the observation of human types.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Data and scripts}\label{scripts}
Messages were downloaded from the Gmane public database~\cite{GMANEwikipedia}.
All routines necessary to achieve the results reported in this article, including tables and figures of Supporting information, are available through a public domain Python package and an open Git repository~\cite{gmanePack}.

>>>>

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in interaction networks. The activity along time and topology evolution were investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparying with the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. A clear distinction could be made of networks with few, stable agents, which possess a large number of threads per message, from others with many agents exhibiting intermittent activity and lower number of threads per message. Also important was the observation that the activity along time, in terms of seconds, minutes, hours, days and months, is practically the same for all lists. These results are coherent with literature and were found also in the analysis of other networks from other emails lists and from Facebook and Twitter. Therefore, the properties observed may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{newmanBook}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{gmane,barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed ~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.

In this paper we analyze the evolution of human interaction networks, by considering interaction in email lists as their representative. Using a timeline of activity snapshots with a constant number of contiguous messages in email lists, we found a remarkable stability for several of the network properties. Because this stability was shared by all email lists, is coherent with literature, and has been observed in some networks from Twitter, Facebook and Participa.br, we advocate that some of the conclusions might be valid for more general classes of interaction networks. In particular, this allows us to discuss typologies in the context of such networks, in an attempt to bridge the gap between approaches based solely on data analysis (i.e. from a hard sciences perspective) and those relevant to the social sciences. This is important insofar as typologies are the canon of scientific literature for classification of human agents~\cite{typCanon}.

The paper is organized as follows. Section~\ref{sec:related} describes related work, while details of the data and methods of analysis are given in  Section~\ref{sec:data} and Section~\ref{sec:carac}. Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions. Appendix has directions to access data and computer scripts to reproduce the analysis. Support Information document has a throughout exposition of measurements in networks from four email lists and a brief analysis of networks from Twitter, Facebook and Participa.br.

@@

We analyzed many email lists (and data from Twitter, Facebook and Participa.br) but selected only four in order to make a thorough analysis, from which general properties can be inferred. These lists, selected as representing both a diverse set and ordinary lists, are:
\begin{itemize}
	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries. English is the language used the most. Abbreviated as LAU from now on.
	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants from different countries, and English is the language used the most. A more technical and less active version of LAU. Abbreviated LAD from now on.
	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries. English is the language used the most. Abbreviated as CPP from now on.
	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests. Portuguese is the most used language, although Spanish and English are also incident. Abbreviated MET from now on.
\end{itemize}

@@

A systematic study of the activity of participants belonging to the three distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices, while the network structure was governed by the intermediary vertices. Because these properties were shared by all email lists and were time-independent, are coherent with literature, and was also found in networks from Facebook, Twitter and Participa.br, we may consider the classification of agents into Erd\"os sectors as leading to a human typology which bridges between exact sciences, with objective procedures for the classification, and human sciences, where there is a legacy in the observation of human types.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section*{APPENDIX: Data and scripts}\label{scripts}
Messages were downloaded from the Gmane public database~\cite{GMANEwikipedia}.
All routines necessary to achieve the results reported in this article, including tables and figures of Supporting information, are available through a public domain Python package and an open Git repository~\cite{gmanePack}.
